---
title: "Using RMarkdown"
teaching: 10
exercises: 2
---

:::::::::::::::::::::::::::::::::::::: questions 

- How do you write a lesson using RMarkdown and `{sandpaper}`?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Explain how to use markdown with the new lesson template
- Demonstrate how to include pieces of code, figures, and nested challenge blocks

::::::::::::::::::::::::::::::::::::::::::::::::

## Introduction


:::::::: callout

## Non-standard packages!

We can now include non-standard packages in our {sandpaper} sites and have them
be reproducible! Case in point: [the {cowsay} 
package](https://cran.r-project.org/package=cowsay) is not one of the imports
for sandpaper, but I can use it to generate random cat facts for my lesson, as
shown below.

### Demo

```{r wisdom, message = FALSE}
library("cowsay")
library("curl")
say("catfact")
```

### Try it yourself!

To try this out, install the development version of sandpaper from my universe:

```{r eval = FALSE}
options(repos = c(
    zkamvar = 'https://zkamvar.r-universe.dev',
    CRAN = 'https://cloud.r-project.org'))

install.packages('sandpaper')
```

Then, clone this repository and run `sandpaper::build_lesson()` inside the 
repository from your R console. Should this work, the dependencies will 
automatically install to a library specific to the lesson and the lesson will
render, regardless if you have {cowsay} previously installed.

:::


This is the new Carpentries template. It is written in [RMarkdown][r-markdown],
which is a variant of Markdown that allows you to render code inside the
lesson. Please refer to the [lesson
example](https://carpentries.github.io/lesson-example) for full documentation.

What you need to know is that there are three block quotes required for a valid
Carpentries lesson template:

 1. `questions` are displayed at the beginning of the episode to prime the
    learner for the content.
 2. `objectives` are the learning objectives for an episode displayed with
    the questions.
 3. `keypoints` are displayed at the end of the episode to reinforce the
    objectives.

::::::::::::::::::::::::::::::::::::: challenge 

## Challenge 1: Can you do it?

What is the output of this command?

```{r, eval = FALSE}
paste("This", "new", "template", "looks", "good")
```

:::::::::::::::::::::::: solution 

## Output
 
```{r, echo = FALSE}
paste("This", "new", "template", "looks", "good")
```

:::::::::::::::::::::::::::::::::


## Challenge 2: how do you nest solutions within challenge blocks?

:::::::::::::::::::::::: solution 

You can add a line with at least three colons and a `solution` tag.

:::::::::::::::::::::::::::::::::
::::::::::::::::::::::::::::::::::::::::::::::::

## Figures

You can also include figures generated from RMarkdown:

```{r pyramid, fig.alt = "pie chart illusion of a pyramid", fig.cap = "Sun arise each and every morning"}
pie(
  c(Sky = 78, "Sunny side of pyramid" = 17, "Shady side of pyramid" = 5), 
  init.angle = 315, 
  col = c("deepskyblue", "yellow", "yellow3"), 
  border = FALSE
)
```

Or you can use standard markdown for static figures:

![You belong in The Carpentries!](https://raw.githubusercontent.com/carpentries/logo/master/Badge_Carpentries.svg){alt='Blue Carpentries hex person logo with no text.'}


## Math

One of our episodes contains $\LaTeX$ equations when describing how to create
dynamic reports with {knitr}, so we now use mathjax to describe this:

`$\alpha = \dfrac{1}{(1 - \beta)^2}$` becomes: $\alpha = \dfrac{1}{(1 - \beta)^2}$

Cool, right?

::::::::::::::::::::::::::::::::::::: keypoints 

- Use `.Rmd` files for lessons even if you don't need to generate any code
- Run `sandpaper::check_lesson()` to identify any issues with your lesson
- Run `sandpaper::build_lesson()` to preview your lesson locally

::::::::::::::::::::::::::::::::::::::::::::::::

```{r sessioninfo, R.options = list(width = 100)}
sessioninfo::session_info()
```


