---
title: "Example with the {poppr} package"
teaching: 10
exercises: 2
---

:::::::::::::::::::::::::::::::::::::: questions 

- Does this example work if you have never used the package before?

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- Introduce the {poppr} package
- Demonstrate resampling of $\bar{r}_d$

::::::::::::::::::::::::::::::::::::::::::::::::

## Introduction

The standardized index of association ($\bar{r}_d$) is a method to assess
multilocus linkage disequilibrium in a random sample from a single population.

This can be calculated via the {poppr} package:

```{r ia, message = FALSE}
library("poppr")
library("ggplot2")
theme_set(theme_bw(base_size = 24))
data(Pinf)
print(Pinf)
set.seed(2021-09-21)
Pinf_ia <- ia(Pinf, sample = 999)
print(Pinf_ia)
```

We can also clone-correct the data to remove repeated genotypes, which may 
represent resamples of the same genet. 

```{r ia_cc}
set.seed(2021-09-21)
Pinf_cc <- ia(clonecorrect(Pinf), sample = 999)
print(Pinf_cc)
```

This index has no known distribution, but we can get an idea of the spread by 
resampling it by the unique number of mulitlocus genotypes.

```{r resample, fig.alt = "histogram of the index of association; a solid vertical line represents the observed value of the index of association; a dotted vertical line to the right represents the observed value of the clone corrected data.", fig.cap = "Distribution of the index of assocation for 999 samples of 72 genotypes"}
# Calculate the resampled index
set.seed(2021-09-21)
Pinf.resamp <- resample.ia(Pinf, reps = 999)

# Create the plot
library("ggplot2")
ggplot(Pinf.resamp[2], aes(x = rbarD)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = Pinf_ia[3]) +
  annotate("label", x = Pinf_ia[3], y = 60, label = "Observed Index", hjust = 0) +
  geom_vline(xintercept = Pinf_cc[3], linetype = 2) +
  annotate("label", x = Pinf_cc[3], y = 50, label = "Observed Index (clone-corrected)", hjust = 0) +
  xlab(expression(bar(r)[d])) +
  xlim(c(0, NA))
```


::::::::::::::::::::::::::::::::::::: keypoints 

 - This is a demo of the {poppr} package

::::::::::::::::::::::::::::::::::::::::::::::::

